---
title: "shannon_r_assessment"
output: html_document
---

**Academic Integrity Notice**

> Your submission must be your own work. You should be prepared to explain any code or answers you submit if asked.  
> Use of generative AI tools (such as ChatGPT, Copilot, Gemini, etc.) to create or modify code or text for this assessment is not permitted.  
> If you consulted any external resources (such as documentation or tutorials), you must include the full URL of each site you used with your answer.

*Datasets*
Two datasets are provided with this assessment:
- emissions tracker data (`climate.csv`)
- a table of country names and ISO country codes (`country_codes.csv`)

The data from climate.csv comes from the [Climate Action Tracker](https://climateactiontracker.org/), an organization that keeps track of countries' progress in tackling climate change. 

*Instructions*
# 1. Create a new R notebook. Name your file as <firstname_R_assessment>.Rmd. For example, my notebook would be named `michael_R_assessment.Rmd`
```{r}
library(tidyverse)
library(janitor)
```


# 2. Read in the data from climate.csv. How many rows and columns does this dataset have?
```{r}
climate_df <- read_csv("../data/climate.csv")

climate_df |> summary()
```
ANSWER: CLIMATE DF CONTAINS 37625 ROWS & 7 COLUMNS


# 3. Read in the country_codes.csv data. This contains country names and 2 and 3-letter ISO country codes.
```{r}
country_codes_df <- read_csv("../data/country_codes.csv")
```


Merge this with the original dataset (keeping only those rows that have matches on the 2-letter country code in both tables). At this point, feel free to rename columns and/or drop any columns that you think will not be useful.
```{r}
climate_country_df <- country_codes_df |> 
    inner_join(
        climate_df,
        by = join_by(`A2 (ISO)` == Country)
    ) |> 
    clean_names(case = "screaming_snake") |> 
    select(COUNTRY, INDICATOR, UNIT, YEAR, VALUE)

climate_country_df |> head(5)
```


# 4. After the merge, how many countries are left in the dataset?
```{r}
climate_country_df |> 
    distinct(COUNTRY) |> 
    count()
```
ANSWER: AFTER THE MERGE, 35 COUNTRIES ARE LEFT IN THE DF


# 5. Create a new dataframe, called "renewable" which contains all rows with indicator "Share of renewable electricity generation" from the **year 2014**.
```{r}
renewable_df <- climate_country_df |> 
    filter(
        INDICATOR == "Share of renewable electricity generation" 
        & YEAR == 2014
    )

renewable_df |> head(5)
```


- Create a visualization (your choice) showing the distribution of the share of renewable electricity generation across the countries in the dataset. 
```{r}
renewable_df |> 
    ggplot(aes(x = VALUE, y = reorder(COUNTRY, VALUE))) +
    geom_col() +
    labs(
        title = "Share of Renewable Electricity Generation",
        subtitle = "2014",
        x = "% Unit",
        y = "Country"
    )
```


```{r}
renewable_df |> 
    ggplot(aes(x = VALUE)) +
    geom_histogram(bins = 8) +
    labs(
        title = "Share of Renewable Electricity Generation",
        subtitle = "2014",
        x = "% Unit",
        y = "Number of Countries"
    )
```



```{r}
renewable_df |> 
    ggplot(aes(x = UNIT, y = VALUE)) +
    geom_boxplot() +
    stat_summary(fun=mean, geom="point", color = "red") +
    labs(
        title = "Share of Renewable Electricity Generation",
        subtitle = "2014",
        x = "",
        y = "% Unit"
    )
```


- What can you say about this distribution? 
ANSWER: THE RENEWABLE ELECTRICITY GENERATION SPANS FROM 0% TO 100% WITH THE MAJORITY UNDER 50% AND THE MEAN BELOW 35%

- Create a new column in renewable, called "majority_renewable" and in this column indicate whether a country is majority renewable (>50% renewable electricity generation) or not. 
```{r}
renewable_df <- renewable_df |> 
    mutate(MAJORITY_RENEWABLE = if_else(VALUE > 50.0, TRUE, FALSE))

renewable_df |> head(5)
```


- What percentage of countries in this data frame were majority renewable in 2014?
```{r}
renewable_df |>
    summarise(sum(MAJORITY_RENEWABLE)) / renewable_df |>
    count()
```
ANSWER: 29.032% OF COUNTRIES WERE MAJORITY RENEWABLE (>50%) IN 2014


- How does the "Electricity emissions intensity" in 2014 compare for countries which were majority renewable in 2014 vs those which were not? Create a visualization (your choice) to support your assessment.
```{r}
renewable_df |> 
    select(COUNTRY, MAJORITY_RENEWABLE) |> 
    right_join(
        climate_country_df |> 
            filter(
                INDICATOR == "Electricity emissions intensity"
                & YEAR == 2014
            )
    ) |> 
    ggplot(aes(x = MAJORITY_RENEWABLE, y = VALUE)) +
    geom_boxplot() +
    labs(
        title = "Electricity Emissions Intensity",
        x = "Majority Renewable Energy Country",
        y = "Electricity Emissions Intensity"
    )
```


# 6. Which country was the top oil and gas producer in 2014? (Hint: this value can be found by using the indicator "Oil and Gas activity: production")
```{r}
oil_gas_df <- climate_country_df |> 
    filter(INDICATOR == "Oil and Gas activity: production" & YEAR == 2014)

oil_gas_df |> 
    slice_max(VALUE)
```
ANSWER: THE UNITED STATES WAS THE TOP OIL AND GAS PRODUCER IN 2014


# 7. Look at the trend in oil and gas production (again, using the "Oil and Gas activity: production" indicator) in the US from 1990 through 2014.
```{r}
us_oil_gas_df <- climate_country_df |> 
    filter(
        INDICATOR == "Oil and Gas activity: production"
        & COUNTRY == "United States"
        & between(YEAR, 1990, 2014)
    )

us_oil_gas_df |> head(5)
```


What do you notice? Create a visualization (your choice) to support your assessment.
```{r}
us_oil_gas_df |> 
    ggplot(aes(x = YEAR, y = VALUE)) +
    geom_line() +
    labs(
        title = "Oil and Gas activity: production",
        subtitle = "United States",
        x = "Year",
        y = "Mtoe Units"
    )
```
ANSWER: THE UNITED STATES OIL AND GAS PRODUCTION EXPONENTIALLY INCREASED AFTER 2010


# 8. Find the total number of new EV sales per year **from 2009 through 2014**. (Sales per country can be found using indicator "New EV sales" and unit "New EV sales".) 
```{r}
ev_sales_df <- climate_country_df |> 
    filter(
        INDICATOR == "New EV sales"
        & UNIT == "New EV sales"
        & between(YEAR, 2009, 2014)
    ) |> 
    group_by(YEAR) |> 
    summarise(TOTAL_SALES = sum(VALUE))

ev_sales_df
```

What do you notice? Create a visualization (your choice) to support your answer.
```{r}
ev_sales_df |> 
    ggplot(aes(x = YEAR, y = TOTAL_SALES)) +
    geom_line() +
    labs(
        title = "New EV Sales",
        x = "Year",
        y = "Sold Units"
    )
```
ANSWER: THERE IS A POSITIVE CORRELATION BETWEEN THE TOTAL EV SALES AND THE YEAR


# 9. Create a scatter plot showing Emissions per capita vs GDP per capita for the year 2014. Hint: You can find these values by using indicators "Emissions per capita" and "GDP per capita". Hint #2: You will probably want to subset and then pivot the data in order to create this plot.
```{r}
emissions_gdp_df <- climate_country_df |> 
    filter(
        INDICATOR %in% c("Emissions per capita", "GDP per capita")
        & YEAR == 2014
    ) |> 
    select(!UNIT) |> 
    pivot_wider(
        names_from = INDICATOR,
        values_from = VALUE
    ) |> 
    clean_names(case = "screaming_snake") |> 
    drop_na()

emissions_gdp_df |> head(5)
```


```{r}
emissions_gdp_df |> 
    ggplot(aes(x = GDP_PER_CAPITA, y = EMISSIONS_PER_CAPITA)) +
    geom_point() +
    labs(
        title = "Emissions per capita vs GDP per capita",
        subtitle = "2014",
        x = "GDP Per Capita",
        y = "Emissions Per Capita"
    )
```


# 10. Find the correlation between emissions per capita and GDP per capita for the year 2014 (dropping NAs). How strong of a (linear) relationship is there between these two measures?
```{r}
emissions_gdp_df |> 
    select(EMISSIONS_PER_CAPITA, GDP_PER_CAPITA) |> 
    cor()
```


```{r}
cor.test(emissions_gdp_df$EMISSIONS_PER_CAPITA, emissions_gdp_df$GDP_PER_CAPITA)
```
ANSWER: THERE IS A STRONG LINEAR RELATIONSHIP BETWEEN EMISSIONS AND GDP PER CAPITA THAT IS STATISTICALLY SIGNIFICANT


Once you are finished, email your file to michael.holloway@nashvillesoftwareschool.com.