---
title: "Warmup 12/20/25"
output: html_notebook
---

For this warmup, you've been provided four csv files containing sales data by region for the years 2019 through 2022 (e.g., sales_2019.csv, sales_2020.csv, etc.).
Each file includes columns for region and sales, but no year column.

```{r}
library(tidyverse)
```

Our goal will be to read in and combine all 4 years.

**Part 1:** Start by creating a vector, years, which contains the numbers 2019 through 2022. Hint: you can use the [sequence operator](https://wec.wur.nl/r/swirl/sequences-of-numbers.html) to accomplish this.

```{r}
years <- seq(2019, 2022)
```


**Part 2:**
To construct a filename, you can use the paste0 function. In the Help panel on the right, search for the paste0 function to see how it works. Try pasting together the string "../data/sales_", the number 2019, and then then string ".csv".
```{r}
paste0("../data/sales_", 2019)
```

**Part 3:** Use the [map function with an anonymous function](https://purrr.tidyverse.org/reference/map.html) to create a list containing a tibble for each year.

```{r}
sales <- map(years, \(years) paste0("../data/sales_", years, ".csv")) |> 
  map(read_csv)
```

**Part 4:** We can use the [bind_rows function](https://dplyr.tidyverse.org/reference/bind_rows.html) to combine all of the sales data into a single tibble. When we do this, we want to make sure to be able to tell which year the data came from. 
First, search the Help for the names function. Use this function to set the name of each tibble in the list you created to the year it represents.
Then, use the bind_rows function along with the .id parameter to create a single tibble with a year column.

```{r}
names(sales) <- years
```


```{r}
all_sales <- bind_rows(sales, .id="year")
```

**Part 5:** Which year had the highest total sales?
```{r}
all_sales |> 
  group_by(year) |> 
  summarize(sum_sales = sum(sales)) |> 
  # arrange(desc(sum_sales)) |> 
  slice_max(sum_sales)
```

**Part 6:** Which region had the highest total combined sales across all 4 years?
```{r}
all_sales |> 
  group_by(region) |> 
  summarize(sum_sales = sum(sales)) |> 
  # arrange(desc(sum_sales)) |> 
  slice_max(sum_sales)
```

