---
title: "Warmup 1/10/26"
output: html_notebook
---

```{r}
library(tidyverse)
library(jsonlite)

job_postings <- fromJSON("../data/job_postings.json") |> 
  as_tibble()

job_postings
```

**Part 1: Salary vs. Years of Experience**
Fit a linear regression model with target salary, using years_experience as the predictor variable.

Hint: See [How to Do Linear Regression in R](https://www.datacamp.com/tutorial/linear-regression-R).

```{r}
# Create a linear regression model
job_lm = lm(salary ~ years_experience, data = job_postings)

# Display the coefficients
job_lm
```

**Part 2:**
What does your model estimate for the average salary for jobs requiring 0 years of experience? 5 years of experience?

Hint: You can use the predict function to generate estimates. You can pass in a tibble for the newdata argument.

Hint 2: You can create a tibble by using the tibble function.

```{r}
# Jobs requiring 0 years experience is equal to the intercept $39,543.

# Create a tibble for years experience to predict
new_data = tibble(years_experience = c(0,5))

# Use predict function with linear regression model and new tibble
predict(object = job_lm, newdata = new_data)
```

**Part 3:** Now, let's see how we can incorporate the skills column into our model.

First, use the unnest function to expand out the skills column.

```{r}
# Unnest the skills column
job_postings <- job_postings |> 
  unnest(cols = skills)

# Display sample
job_postings |> 
  slice_sample(n = 2)
```

We're going to use the pivot_wider function. To make this work, we need to start by creating a "value" column that contains 1. Use the mutate function to accomplish this.

```{r}
# Create a column 
job_postings <- job_postings |> 
  mutate(value = 1, .after = skills)

# Display sample
job_postings |> 
  slice_sample(n = 2)
```


Now, use the pivot_wider function to create a column for each skill. Make sure to add a prefix of "skill_" to the column names.
```{r}
# Pivot wider to create columns for each skill
job_postings_pivot <- job_postings |> 
  pivot_wider(names_from = skills, values_from = value, names_prefix = "skill_", values_fill = 0)

# Display head
job_postings_pivot |> 
  head(5)
```

Finally, we need to fill in the missing values with 0's. 

Use mutate in combination with the [across function](https://dplyr.tidyverse.org/reference/across.html). Hint: you can select columns inside of across using [starts_with](https://tidyselect.r-lib.org/reference/starts_with.html). 

You can use the coalesce function to fill in missing values with a 0.

```{r}
# Practice using mutate across
job_postings_pivot_2 <- job_postings |> 
  pivot_wider(names_from = skills, values_from = value, names_prefix = "skill_")|> 
  mutate(across(starts_with("skill"), ~replace_na(.x, 0)))

# Display head
job_postings_pivot_2 |> 
  head(5)
```

```{r}
# Practice using mutate across
job_postings_pivot_3 <- job_postings |> 
  pivot_wider(names_from = skills, values_from = value, names_prefix = "skill_")|> 
  mutate(across(starts_with("skill"), ~coalesce(.x, 0)))

# Display head
job_postings_pivot_3 |> 
  head(5)
```

Now, fit a linear regression model for salary using the years_experience plus all of the skill columns.

Which skill has the largest positive impact on the estimated average salary?

```{r}
job_lm_2 = lm(salary ~ ., data = job_postings_pivot |> select(!title))

job_lm_2
```
