---
title: "Warmup 01/15/26"
output: html_notebook
---

In this warmup, you'll see how to build a logistic regression model using R.

For this, we'll be using the Titanic dataset.

```{r}
library(tidyverse)

titanic <- read_csv("../data/titanic.csv")
titanic
```

**Question 1:** Fit a logistic regression model to estimate the probability that a passenger survived based on their age.

Hint: Check out https://statsandr.com/blog/binary-logistic-regression-in-r/#binary-logistic-regression-in-r#binary-logistic-regression-in-r to see the syntax for fitting a logistic regression model in R.

```{r}
# Create a binary logicstic regression model
model = glm(Survived ~ Age, data = titanic, family = "binomial")

summary(model)
```
```{r}
# Get the odds ratio for Age
exp(coef(model)["Age"])
```

What does the model estimate for the probability of survival for a 20 year old passenger? For a 70 year old passenger?

Hint: You can use the predict function, like you did for linear regression. However, you'll need to specify the type of output you want. See https://www.statology.org/r-logistic-regression-predict/
```{r}
# Calculate probability of survival for 20 years old
predict(model,
  newdata = data.frame(Age = c(20, 70)),
  type = "response"
)
```
```{r}
# Calculate probability of survival for 70 years old
predict(model,
  newdata = data.frame(Age = c(70)),
  type = "response"
)
```

**Part 2:** Fit a logistic regression model to estimate the probability of survival based on the passenger's sex.
```{r}
# Create a binary logicstic regression model
model = glm(Survived ~ Sex, data = titanic, family = "binomial")

summary(model)
```
```{r}
# Get the odds ratio for Sex
exp(coef(model)["Sex"])
```

What does the model estimate as the probability of survival for a male passenger? For a female passenger?
```{r}
# Calculate probability of survival for male
predict(model,
  newdata = data.frame(Sex = c("male", "female")),
  type = "response"
)
```
```{r}
# Calculate probability of survival for female
predict(model,
  newdata = data.frame(Sex = c("female")),
  type = "response"
)
```

**Part 3:** Fit a new model to estimate the probability of survival using age, sex, and passenger class.
Make sure that the model is treating the passenger class as a categorical variable, not as a numeric one.
Hint: You may want to convert the Pclass variable to a factor. See https://www.statology.org/r-linear-regression-with-categorical-variables/
You can read more about factors in [Chapter 15 of R for Data Science](https://r4ds.had.co.nz/factors.html).


```{r}
# Convert Pclass to a factor type
titanic$Pclass <- as.factor(titanic$Pclass)
```


```{r}
# Create a binary logicstic regression model
model = glm(Survived ~ Age + Sex + Pclass, data = titanic, family = "binomial")

summary(model)
```

What does the model estimate for the probability of a 20 year old female in first class surviving? A 50 year old male in 3rd class?

```{r}
# Calculate probability of survival for 20 year old female in first class
predict(model,
  newdata = data.frame(Age = c(20, 50), Sex = c("female", "male"), Pclass = c("1", "3")),
  type = "response"
)
```

```{r}
# Calculate probability of survival for 50 year old male in 3rd class
predict(model,
  newdata = data.frame(Age = c(50), Sex = c("male"), Pclass = c("3")),
  type = "response"
)
```

