---
title: "Warmup 26_01_24"
output: html_notebook
---

```{r}
library(tidyverse)
```

**Part 1:**
The Public Health Department wants to see whether the number of community outreach events affects the number of people vaccinated in different districts.

The following tibble contains the number of outreach events and the number vaccinated by district.

```{r}
vaccinations <- tibble(
  district = c("A", "B", "C", "D", "E"),
  outreach_events = c(2, 5, 3, 7, 4),
  vaccinated = c(40, 120, 65, 200, 90)
)
vaccinations
```

Fit a Poisson regression predicting vaccinated from outreach events.

```{r}

```

Interpret the coefficient you get for outreach_events.
Hint: You can use the coef function to extract the coefficients from a model.

```{r}

```


Estimate the average number vaccinated in a district that holds 2 outreach events. 
Note: you may need to use the type argument ot the predict function. See https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/predict.glm 

```{r}

```


**Part 2:**

The department now wants to look at reported cases of respiratory illness across counties with different population sizes, trying to see the effect of a county being urban or rural.

```{r}
cases <- tibble(
  county = c("X", "Y", "Z", "W", "V", "U"),
  cases = c(80, 30, 100, 25, 110, 20),
  population = c(120000, 80000, 200000, 70000, 250000, 60000),
  urban = c(1, 0, 1, 0, 1, 0)
)
cases
```

Why would modeling cases alone be misleading?

Fit a Poisson regression model estimated the average number of cases based on the urban variable, using population as an offset.

```{r}

```

Interpret the coefficient on urban.

```{r}

```

What does the model estimate for the average number of cases for an urban county with a population of 150000? For a rural county with a population of 40000?

```{r}

```
