---
title: "Using Postgre with R"
output: html_notebook
---

First, we need to establish a connection with our database. For this, we will use the RPostgres and DBI libraries.

If you are working with a large database, you might consider using the dbplyr library (https://dbplyr.tidyverse.org/articles/dbplyr.html), which allows you to use dplyr verbs to interact with tables in a database, but actually execute them in the database.

```{r}
library(DBI)
library(RPostgres)
library(tidyverse)
```

Now, we'll create our connection.
```{r}
con <- dbConnect(Postgres(),                 # Use the postgres driver
                 dbname = 'prescribers',     # Name of the database we want to connect to
                 host = 'localhost',         # 'localhost' or eg. 'ec2-54-83-201-96.compute-1.amazonaws.com'
                 port = 5432, 
                 user = 'postgres',
                 password = rstudioapi::askForPassword("Database password"))
```

We can see the tables in the database.
```{r}
dbListTables(con)
```

We can even see the columns of a particular table.
```{r}
dbListFields(con, 'drug')
```

To execute a query, we can use the dbSendQuery function
```{r}
query = "SELECT * FROM prescriber LIMIT 100;"

res <- dbSendQuery(con, query)
```

Then to retrieve the results, we can use dbFetch. Note that you can fetch only a portion of the results if needed, but we'll fetch everything.
```{r}
prescribers <- dbFetch(res)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(res)
```

The results that are returned come through as a dataframe, so we can manipulate it using tidyverse.
```{r}
prescribers %>% 
  filter(nppes_provider_city == 'NASHVILLE')
```